---
title: "Grace Mao Allen Shi Final Project"
output: rmarkdown::html_document
runtime: shiny
---

<style type="text/css">
.table {
    width: 40%;
    position: relative;
    left: 7.5%;
}
</style>

<br/><br/>

### Summary

We got json files from ratemyprofessor
Our goal was to find interesting trends
looked at Duke and UNC

<br/>


#### Data Collection

summarize here

```{r}
library(rvest)
library(httr)
library(XML)
library(jsonlite)

get_data <- function(school){
  final <- data.frame(NULL)
  
  #we take the first 80 pages for each school
  for (i in 1:80){
    url <- paste0("http://www.ratemyprofessors.com/find/professor/?department&institution=Duke%20University&page=",i,
                  "&query&queryoption=TEACHER&queryBy=schoolId&sid=",school,"&sortBy")

    data<-fromJSON(url)

    dept = NULL
    first_name = NULL
    last_name = NULL
    num_rating = NULL
    rating_class = NULL
    overall_rating = NULL
    
    #there are 20 teacher entries for each page 
    for(k in 1:20){
      dept[k]           = data$professors[[k]][[1]]
      first_name[k]     = data$professors[[k]][[4]]
      last_name[k]      = data$professors[[k]][[6]]
      num_rating[k]     = strtoi(data$professors[[k]][[8]])
      rating_class[k]   = data$professors[[k]][[9]]
      overall_rating[k] = as.numeric(data$professors[[k]][[12]])
    }
    
    #make data frame for page
    temp = data.frame(dept, first_name, last_name, num_rating, rating_class, overall_rating, stringsAsFactors = FALSE)
    #add these rows to total school data frame
    final <- rbind(final, temp)
  }
  #return school data frame 
  return (final)
}

#make data frames for each school
school_data <- lapply(list(Duke=1350, UNC=1232), function(x) get_data(x))
```

### Data cleanup

look at sorted unique list of department names for Duke and UNC to determine if some department categories should be combined

```{r}
school_data$Duke[school_data$Duke$dept == "Art History", ]$dept <- "Art & Art History"
school_data$Duke[school_data$Duke$dept == "Art  Art History", ]$dept <- "Art & Art History"
school_data$Duke[school_data$Duke$dept == "Physical Ed", ]$dept <- "Physical Education"
school_data$Duke[school_data$Duke$dept == "Athletics", ]$dept <- "Physical Education"
school_data$Duke[school_data$Duke$dept == "Women's Studies", ]$dept <- "Womens Studies"

school_data$UNC[school_data$UNC$dept == "Environmental Science", ]$dept <- "Environmental Sci & Eng"
school_data$UNC[school_data$UNC$dept == "Biological Sciences", ]$dept <- "Biology"
school_data$UNC[school_data$UNC$dept == "Physical Ed", ]$dept <- "Physical Education"
school_data$UNC[school_data$UNC$dept == "Exercise  Sport Science", ]$dept <- "Exercise & Sport Science"
school_data$UNC[school_data$UNC$dept == "Art", ]$dept <- "Art & Art History"
school_data$UNC[school_data$UNC$dept == "Art History", ]$dept <- "Art & Art History"
school_data$UNC[school_data$UNC$dept == "Health Policy & Management", ]$dept <- "Health Administration"
school_data$UNC[school_data$UNC$dept == "Piano", ]$dept <- "Music"

```

### Data manipulation

average rating per department
multiply overall rating by number of ratings to get total rating, then divide by total number of ratings
the trend is

```{r}

duke_dept_avg <- lapply(unique(school_data$Duke$dept), 
                      function(x) 
                        sum(school_data$Duke[school_data$Duke$dept == x, ]$overall_rating * 
                        school_data$Duke[school_data$Duke$dept == x, ]$num_rating)/
                        sum(school_data$Duke[school_data$Duke$dept == x, ]$num_rating))
names(duke_dept_avg) <- unique(school_data$Duke$dept)
duke_dept_avg <- sort(unlist(duke_dept_avg),decreasing=TRUE)
duke_dept_avg

unc_dept_avg <- lapply(unique(school_data$UNC$dept), 
                     function(x) 
                       sum(school_data$UNC[school_data$UNC$dept == x, ]$overall_rating * 
                             school_data$UNC[school_data$UNC$dept == x, ]$num_rating)/
                       sum(school_data$UNC[school_data$UNC$dept == x, ]$num_rating))
names(unc_dept_avg) <- unique(school_data$UNC$dept)
unc_dept_avg<-sort(unlist(unc_dept_avg),decreasing=TRUE)
unc_dept_avg
```

number of ratings per department
the trend is  

```{r}
duke_num_ratings <- lapply(unique(school_data$Duke$dept), 
                      function(x) 
                        sum(school_data$Duke[school_data$Duke$dept == x, ]$num_rating))
names(duke_num_ratings) <- unique(school_data$Duke$dept)
duke_num_ratings <- sort(unlist(duke_num_ratings),decreasing=TRUE)
duke_num_ratings

unc_num_ratings <- lapply(unique(school_data$UNC$dept), 
                           function(x) 
                             sum(school_data$UNC[school_data$UNC$dept == x, ]$num_rating))
names(unc_num_ratings) <- unique(school_data$UNC$dept)
unc_num_ratings <- sort(unlist(unc_num_ratings),decreasing=TRUE)
unc_num_ratings
```

Find number of ratings per rating class
the trend is that students tend to give more ratings to the "better" professors

```{r}
duke_class_num <- lapply(unique(school_data$Duke$rating_class), 
                           function(x) 
                             sum(school_data$Duke[school_data$Duke$rating_class == x, ]$num_rating))
names(duke_class_num) <- unique(school_data$Duke$rating_class)
duke_class_num <- sort(unlist(duke_class_num),decreasing=TRUE)
duke_class_num

unc_class_num <- lapply(unique(school_data$UNC$rating_class), 
                         function(x) 
                           sum(school_data$UNC[school_data$UNC$rating_class == x, ]$num_rating))
names(unc_class_num) <- unique(school_data$UNC$rating_class)
unc_class_num <- sort(unlist(unc_class_num),decreasing=TRUE)
unc_class_num

```

out of the professors with at least 30 ratings, sort by best rated
looks like best rated profs at Duke are STEM classes and best rated profs at UNC are history/humanities

```{r}
valid_duke_profs <- school_data$Duke[school_data$Duke$num_rating >= 30, ]
valid_unc_profs <- school_data$UNC[school_data$UNC$num_rating >= 30, ]
valid_duke_profs[order(-valid_duke_profs[,6]), ][1:10,]
valid_unc_profs[order(-valid_unc_profs[,6]), ][1:10,]
```

what is the most common first name of professors lol
top 10 are basically the same for each school

```{r}
dukenames<-count(school_data$Duke, 'first_name')
dukenames[order(-dukenames[,2]), ][1:10,]
uncnames<-count(school_data$UNC, 'first_name')
uncnames[order(-uncnames[,2]), ][1:10,]
```

### Testing
make some comparisons between Duke and UNC with tests or just comparing the above

```{r}
t.test(school_data$Duke$overall_rating, school_data$UNC$overall_rating)
t.test(school_data$Duke$num_rating, school_data$UNC$num_rating) #hey this is significant!
```

